---
title: "Wiley Approach"
author: "Najko Jahn"
date: "10/28/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Required R libraries

```{r}
library(tidyverse) # collection of packages, see also <https://r4ds.had.co.nz/>
library(jsonlite) # json tools to import and export nested data
library(rcrossref) # interface to Crossref API
library(stringi) # string manipulation
```

## Journal information

The starting point for the ESAC market watch report is to obtain information about the journal portfolio of a publisher. Many publishers provide detailled information about their portfolio. 

Information about Wiley open access options can be found here:

<https://authorservices.wiley.com/author-resources/Journal-Authors/open-access/article-publication-charges.html>

- fully open access <https://authorservices.wiley.com/asset/Wiley-Journal-APCs-Open-Access.xlsx>
- hybrid open access <https://authorservices.wiley.com/asset/Wiley-Journal-APCs-OnlineOpen.xlsx>


Download the files:

```{r}
download.file(url = "https://authorservices.wiley.com/asset/Wiley-Journal-APCs-Open-Access.xlsx", 
              destfile = "data/wiley_full_oa_journal_list.xlsx")

download.file(url = "https://authorservices.wiley.com/asset/Wiley-Journal-APCs-OnlineOpen.xlsx", 
              destfile = "data/wiley_hybrid_oa_journal_list.xlsx")
```

Prepare the journal list:

```{r}
full_oa <- readxl::read_xlsx("data/wiley_full_oa_journal_list.xlsx", skip = 3)
full_oa_short <- full_oa %>%
  select(journal_title = Journal, jn_code = `Journal Code`, e_issn = `Online ISSN`) %>%
  mutate(oa_mode = "full oa") %>%
  filter(!is.na(journal_title)) 
```

```{r}
hybrid_oa <- readxl::read_xlsx("data/wiley_hybrid_oa_journal_list.xlsx", skip = 3)
hybrid_oa_short <- hybrid_oa %>% 
  select(journal_title = `Journal Title`, jn_code = `Journal Code`, e_issn = `Online\r\nISSN`) %>%
  mutate(oa_mode = "hybrid oa") %>%
  filter(!is.na(jn_code)) 
jn_wiley <- bind_rows(hybrid_oa_short, full_oa_short)
```

Wiley journal list contains just Online ISSN. To increase the recall when calling the Crossref API by ISSN, we add other ISSN variants to the journal dataset from the ISSN-L journal list.

<https://www.issn.org/services/online-services/access-to-issn-l-table/>

Download: <http://www.issn.org/wp-content/uploads/2014/03/issnltables.zip>

```{r}
issn_l <- readr::read_tsv("data/20200119.ISSN-to-ISSN-L.txt")
issn_match <- jn_wiley %>%
  left_join(issn_l, by = c("e_issn" = "ISSN")) 
issn_variants <-
  issn_match %>% 
  left_join(issn_l, by = c("ISSN-L")) %>%
  distinct(jn_code, e_issn, `ISSN-L`, ISSN) %>%
  gather(e_issn, `ISSN-L`, ISSN, key = "issn_type", value = issn) %>%
  distinct()
```

Per every ISSN-L, obtain all ISSN variants and prepare Crossref calls


### Publication volume per journal from Crossref

Crossref is a DOI registration agency for scholarly works. It not only mints DOIs for journal articles, but also provides metadata, which makes it a great data source for the ESAC market watch report. 

Its API, which can be accessed with the R package [`rcrossref`](https://cran.r-project.org/package=rcrossref), returns article-level metadata and [facet counts](https://github.com/CrossRef/rest-api-doc#facet-counts) for a specific query. In the following, we will make use of facet counts to obtain 

- the yearly publication volume per journal
- and licence URLs 

by all ISSNs per journal.

First, we create a list wit hqueries that we want to send to Crossref


```{r}
issns_list <-
  purrr::map(unique(issn_variants$jn_code), function(x) {
    jn <- x
    issns <- issn_variants %>%
      filter(jn_code == jn) %>%
      distinct(issn) %>%
      .$issn
    names(issns) <- rep("issn", length(issns))
    issns
  })
```


Next, we call the Crossref API.

**Note that we just query 10 journals for the sake of this demonstration. Calling Crossref API for around 2000 journals takes around 1-2 hours if you have registered for the polite pool: <https://github.com/ropensci/rcrossref#register-for-the-polite-pool>**

```{r}
jn_facets <- purrr::map(issns_list[c(991:1000)], .f = purrr::safely(function(x) {
  tt <- rcrossref::cr_works(
    filter = c(
      x,
      from_pub_date = "2015-01-01",
      until_pub_date = "2019-12-31",
      type = "journal-article"
    ),
    # being explicit about facets improves API performance!
    facet = "license:*,published:*",
    # less api traffic
    select = "DOI"
  )
  #' Parse the relevant information
  #' - `issn` - issns  found in open apc data set
  #' - `year_published` - published volume per year (Earliest year of publication)
  #' - `license_refs` - facet counts for license URIs of work
  if (!is.null(tt)) {
    tibble::tibble(
      issn = list(x),
      year_published = list(tt$facets$published),
      license_refs = list(tt$facets$license)
    )
  } else {
    NULL
  }
}))
```

`purrr::safely()` was used to catch potential errors. To get the publication volume:

```{r}
jn_df <- jn_facets %>%
  map_df("result")
```

Publication volume

```{r}
pub_volume <- jn_df %>% 
  select(-license_refs) %>%
  unnest(year_published) %>% 
  unnest(issn) %>%
  inner_join(issn_variants, by = "issn") %>%
  select(year = .id, articles = V1, jn_code) %>%
  distinct() %>%
  inner_join(jn_wiley, by = "jn_code")
pub_volume
```

### Open Access articles

While we can infer that all articles from fully open access journal are openly available, identifying open access articles in subscription journals, i.e. hybrid open access, requires another step. 

Hybrid open access articles are made available under an open content license like Creative Commons.

First, we check, if a subscription journal publsihed articles under an Creative Commons license

```{r}
license_jns <- jn_df %>% 
  select(-year_published) %>%
  unnest(license_refs) %>% 
  unnest(issn) %>%
    inner_join(issn_variants, by = "issn") %>%
  select(license_ref = .id, articles = V1, jn_code, issn) %>%
  distinct() %>%
  inner_join(jn_wiley, by = "jn_code") %>%
  mutate(hybrid_license = ifelse(grepl(
    "creativecommons",
    license_ref
  ), TRUE, FALSE))  %>%
  filter(hybrid_license == TRUE, oa_mode == "hybrid oa") %>%
  nest(issn = c(issn))
license_jns 
```

We now know, whether and which open licenses were used by the journal.
As a next step we want to validate that these licenses were not issued for delayed open access articles by additionally using  the self-explanatory filter `license.url` and `license.delay`. We also obtain metadata for these hybrid open access articles stored as list-column. Parsed metadata fields are defined in `cr_md_fields`

```{r}
cr_md_fields <- c("URL", "member", "created", "license", 
                   "container-title", "issued", "approved", 
                  "indexed", "accepted", "DOI", "funder", "published-print", 
                  "subject", "published-online", "link", "type", "publisher", 
                  "issn-type", "deposited", "content-created")
cr_license <- purrr::map2(license_jns$license_ref, license_jns$issn,
                          .f = purrr::safely(function(x, y) {
                            u <- x
                            issn <- y
                            names(issn) <-rep("issn", length(issn))
                            tmp <- rcrossref::cr_works(filter = c(issn, 
                                                                  license.url = u, 
                                                                  type = "journal-article",
                                                                  from_pub_date = "2015-01-01", 
                                                                  until_pub_date = "2019-12-31"),
                                                       cursor = "*", cursor_max = 5000L, 
                                                       limit = 1000L,
                                                       select = cr_md_fields) 
                            tibble::tibble(
                              issns =  list(issn),
                              license_ref = u,
                              md = list(tmp$data)
                            )
                          }))
#' into one data frame!
cr_license_df <- cr_license %>% 
  purrr::map_df("result") 
#' export results as nested json. For re-importing it into R, use the stream_in function
dplyr::bind_rows(cr_license_df) %>% 
  jsonlite::stream_out(file("data/hybrid_license_md.json"))
```

Exclude delayed OA

```{r}
#' get delayed oa  articles (<= 31 days)
immediate_dois <- cr_license_df %>%
  unnest(md) %>% 
  select(doi, license) %>% 
  unnest(license) %>% 
  filter(grepl("creativecommons", URL), delay.in.days <= 31)
```

Some data transformation to obtain hybrid oa figures per journal and year from the article-level data.

```{r}
license_df <- cr_license_df %>% 
  unnest(md) %>% 
  unnest(issns) %>%
  inner_join((license_jns %>% unnest(issn)), by = "issn") %>%
  mutate(year = str_extract(issued, "[0-9]{4}")) %>%
  group_by(year, jn_code) %>%
  summarise(hybrid_oa_articles = n_distinct(doi))
license_df
```

Finally, add the overall publication volume

```{r}
inner_join(license_df, pub_volume, by = c("jn_code", "year"))
```